|

### 2023-01-23 paper_comparison/appendix.Rmd
combine two different x_dist plots into one

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(visage)
```

# Appendix

## Effect derivation

## $p$-value scatter plot

```{r p-value-comparison-old, eval = FALSE, fig.height = 4, fig.cap = "Visual test p-value compared to conventional test p-value for lineups produced by heteroskedasticity and non-linearity model. The scatter plot is drawn on a square root scale. Every dot on the plot corresponds to a lineup. Most of the dots are fallen on the left of the y = x line indicating visual test p-value is almost always higher than conventional test p-value. The blue curve is a smoothing of the dots showing a positive trend. "}
vi_survey %>%
  filter(!null_lineup) %>%
  filter(!attention_check) %>%
  mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
  left_join(vi_survey %>% 
           count(unique_lineup_id) %>%
           rename(num_subject = n)) %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  mutate(type = factor(type, levels = c("non-linearity", "heteroskedasticity"))) %>%
  ggplot() +
  geom_abline() +
  geom_point(aes(conventional_p_value, p_value), alpha = 0.4) +
  geom_smooth(aes(conventional_p_value, p_value), se = FALSE, span = 0.6) +
  scale_x_sqrt() +
  scale_y_sqrt() +
  facet_wrap(~type) +
  ylab("Visual test p-value on square root scale") +
  xlab("Conventional test p-value on square root scale") +
  theme_light()
```

## A collection interesting lineups (unusual results)

why unusual? what is the possible explanations?

# targe journal

JRSSB: Journal of the Royal Statistical Society Series B (Statistical Methodology)
Deadline: Jan 1, Apr 1

Reading: style of writing, author guideline (https://rss.onlinelibrary.wiley.com/hub/journal/14679868/author-guidelines)

JCGS
JDSS

### 2023-01-23 paper_comparison/paper_comparison/paper_comparison.Rmd
combine two different x_dist plots into one

```{r different-x-dist-poly-power, fig.height = 4, fig.cap = "Comparison of power of conventional tests and visual tests on lineups produced by the non-linearity model under different fitted value distributions."}

dat_different_dist <- poly_conv_sim %>% 
  mutate(log_effect_size = log(effect_size)) %>%
  select(x_dist, log_effect_size, RESET4_p_value) %>%
  rename(RESET_p_value = RESET4_p_value) %>%
  pivot_longer(RESET_p_value) %>%
  mutate(x_dist = ifelse(x_dist == "even_discrete", 
                                       "Discrete", 
                                       stringr::str_to_title(x_dist))) %>%
  mutate(name = gsub("_p_value", " test", name)) %>%
  mutate(reject = value < 0.05) %>%
  mutate(conv_or_not = "Conventional") %>%
  bind_rows(vi_survey %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  filter(type == "polynomial") %>%
  filter(!null_lineup) %>%
  filter(!attention_check) %>%
  mutate(reject = as.numeric(p_value <= 0.05)) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  mutate(x_dist = ifelse(x_dist == "even_discrete", 
                         "Discrete", 
                         stringr::str_to_title(x_dist))) %>%
  mutate(conv_or_not = "Visual")) %>%
  mutate(x_dist = factor(x_dist, 
      levels = c("Uniform", "Normal", 
                 "Lognormal", "Discrete"))) %>%
  select(x_dist, log_effect_size, reject, conv_or_not)

dat_different_dist %>%
  ggplot() +
  geom_smooth(data = ~filter(.x, x_dist == "Uniform"),
              aes(log_effect_size, 
                  reject), 
                  col = "grey", 
              method = "glm", 
              method.args = list(family = binomial), 
              se = FALSE, size = 3) +
  geom_smooth(aes(log_effect_size, 
                  reject, 
                  col = x_dist), 
              method = "glm", 
              method.args = list(family = binomial), 
              se = FALSE, 
              size = 1) +
  theme_light() +
  scale_color_manual(values = rcartocolor::carto_pal(4, "Safe")[c(4,1,3,2)]) +
  scale_size_manual(values = c(1, 2, 3, 4)) +
  xlab(expression(log[e] (Effect_size))) +
  ylab("Power") +
  labs(col = "", linetype = "Test") +
  facet_wrap(~conv_or_not) +
  theme(legend.position="bottom")
```

```{r different-x-dist-heter-power, fig.height = 4, fig.cap = "Power of conventional tests and visual tests on lineups produced by the heteroskedasticity model under different distributions of regressor."}

dat_different_dist <- heter_conv_sim %>% 
  mutate(log_effect_size = log(effect_size)) %>%
  select(x_dist, log_effect_size, BP_p_value) %>%
  pivot_longer(BP_p_value) %>%
  mutate(x_dist = ifelse(x_dist == "even_discrete", 
                                       "Discrete", 
                                       stringr::str_to_title(x_dist))) %>%
  mutate(name = gsub("_p_value", " test", name)) %>%
  mutate(reject = value < 0.05) %>%
  mutate(conv_or_not = "Conventional") %>%
  bind_rows(vi_survey %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  filter(type == "heteroskedasticity") %>%
  filter(!null_lineup) %>%
  filter(!attention_check) %>%
  mutate(reject = as.numeric(p_value <= 0.05)) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  mutate(x_dist = ifelse(x_dist == "even_discrete", 
                         "Discrete", 
                         stringr::str_to_title(x_dist))) %>%
  mutate(conv_or_not = "Visual")) %>%
  mutate(x_dist = factor(x_dist, 
      levels = c("Uniform", "Normal", 
                 "Lognormal", "Discrete"))) %>%
  select(x_dist, log_effect_size, reject, conv_or_not)

dat_different_dist %>%
  ggplot() +
  geom_smooth(data = ~filter(.x, x_dist == "Uniform"),
              aes(log_effect_size, 
                  reject), 
                  col = "grey", 
              method = "glm", 
              method.args = list(family = binomial), 
              se = FALSE, size = 3) +
  geom_smooth(aes(log_effect_size, 
                  reject, 
                  col = x_dist), 
              method = "glm", 
              method.args = list(family = binomial), 
              se = FALSE, 
              size = 1) +
  theme_light() +
  scale_color_manual(values = rcartocolor::carto_pal(4, "Safe")[c(4,1,3,2)]) +
  scale_size_manual(values = c(1, 2, 3, 4)) +
  xlab(expression(log[e] (Effect_size))) +
  ylab("Power") +
  labs(col = "", linetype = "Test") +
  facet_wrap(~conv_or_not) +
  theme(legend.position="bottom")
```

@@

Shape of the non-linearity pattern is another factor contributed to the simulation model.

@@

heter_conv_sim %>%
  filter(x_dist == "uniform") %>%
  mutate(log_effect_size = log(effect_size)) %>%
  mutate(reject = as.numeric(BP_p_value < 0.05)) %>%
  mutate(a = c("Left-triangle", "Butterfly", "Right-triangle")[a + 2]) %>%
  mutate(conv_or_not = "Conventional") %>%
  bind_rows(vi_survey %>%
  filter(!attention_check) %>%
  filter(!null_lineup) %>%
  filter(x_dist == "uniform", type == "heteroskedasticity") %>%
  mutate(log_effect_size = log(effect_size)) %>%
  mutate(reject = as.numeric(p_value < 0.05)) %>%
  mutate(a = c("Left-triangle", "Butterfly", "Right-triangle")[a + 2]) %>%
  mutate(conv_or_not = "Visual")) %>%
  mutate(a = factor(a, levels = c("Left-triangle", "Butterfly", "Right-triangle"))) %>%
  select(a, log_effect_size, reject, conv_or_not) %>%
  ggplot() +
  geom_smooth(aes(log_effect_size, reject, col = a), 
              se = FALSE, 
              method = "glm",
              method.args = list(family = binomial())) +
  xlab(expression(log[e] (Effect_size))) +
  ylab("Power") +
  scale_color_manual(values = rcartocolor::carto_pal(3, "Vivid")) +
  theme(legend.position = "bottom") +
  theme_light() +
  facet_wrap(~conv_or_not) +
  labs(col = "Heteroskedasticity shape")
```

# Appendix

## Effect derivation

## $p$-value scatter plot

```{r p-value-comparison-old, eval = FALSE, fig.height = 4, fig.cap = "Visual test p-value compared to conventional test p-value for lineups produced by heteroskedasticity and non-linearity model. The scatter plot is drawn on a square root scale. Every dot on the plot corresponds to a lineup. Most of the dots are fallen on the left of the y = x line indicating visual test p-value is almost always higher than conventional test p-value. The blue curve is a smoothing of the dots showing a positive trend. "}
vi_survey %>%
  filter(!null_lineup) %>%
  filter(!attention_check) %>%
  mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
  left_join(vi_survey %>% 
           count(unique_lineup_id) %>%
           rename(num_subject = n)) %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  mutate(type = factor(type, levels = c("non-linearity", "heteroskedasticity"))) %>%
  ggplot() +
  geom_abline() +
  geom_point(aes(conventional_p_value, p_value), alpha = 0.4) +
  geom_smooth(aes(conventional_p_value, p_value), se = FALSE, span = 0.6) +
  scale_x_sqrt() +
  scale_y_sqrt() +
  facet_wrap(~type) +
  ylab("Visual test p-value on square root scale") +
  xlab("Conventional test p-value on square root scale") +
  theme_light()
```

## A collection interesting lineups (unusual results)

why unusual? what is the possible explanations?

# targe journal

JRSSB: Journal of the Royal Statistical Society Series B (Statistical Methodology)
Deadline: Jan 1, Apr 1

Reading: style of writing, author guideline (https://rss.onlinelibrary.wiley.com/hub/journal/14679868/author-guidelines)

JCGS
JDSS

>>>>

```{r different-x-dist-poly-power, fig.cap = "Comparison of power of conventional tests and visual tests on lineups produced by the non-linearity model under different fitted value distributions."}

# Extract three datasets, one for poly conventional, one for heter conventional
# and one for visual tests.
list(
poly_conv_sim %>% 
  mutate(log_effect_size = log(effect_size)) %>%
  select(x_dist, log_effect_size, RESET4_p_value) %>%
  rename(RESET_p_value = RESET4_p_value) %>%
  mutate(x_dist = ifelse(x_dist == "even_discrete",
                         "Discrete", 
                         stringr::str_to_title(x_dist))) %>%
  mutate(reject = as.numeric(RESET_p_value < 0.05)) %>%
  mutate(conv_or_not = "Conventional") %>%
  mutate(type = "non-linearity") %>%
  select(type, x_dist, log_effect_size, reject, conv_or_not)
,
heter_conv_sim %>% 
  mutate(log_effect_size = log(effect_size)) %>%
  select(x_dist, log_effect_size, BP_p_value) %>%
  mutate(x_dist = ifelse(x_dist == "even_discrete",
                         "Discrete", 
                         stringr::str_to_title(x_dist))) %>%
  mutate(reject = as.numeric(BP_p_value < 0.05)) %>%
  mutate(conv_or_not = "Conventional") %>%
  mutate(type = "heteroskedasticity") %>%
  select(type, x_dist, log_effect_size, reject, conv_or_not)
,
vi_survey %>%
  filter(!null_lineup) %>%
  filter(!attention_check) %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
  mutate(reject = as.numeric(p_value <= 0.05)) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  mutate(x_dist = ifelse(x_dist == "even_discrete", 
                         "Discrete", 
                         stringr::str_to_title(x_dist))) %>%
  mutate(conv_or_not = "Visual") %>%
  select(type, x_dist, log_effect_size, reject, conv_or_not)
) %>%

  # Bind rows
  map_df(~.x) %>%

  # Manually control factor ordering
  mutate(type = factor(type, 
                       levels = c("non-linearity", 
                                  "heteroskedasticity"))) %>%
  mutate(x_dist = factor(x_dist, 
                         levels = c("Uniform", 
                                    "Normal", 
                                    "Lognormal", 
                                    "Discrete"))) %>%

  ggplot() +

  # Bold the uniform smooth line
  geom_smooth(data = ~filter(.x, x_dist == "Uniform"),
              aes(log_effect_size, reject), 
              col = "grey", 
              method = "glm", 
              method.args = list(family = binomial), 
              se = FALSE, 
              size = 3) +

  # Draw all smooth lines
  geom_smooth(aes(log_effect_size, reject, col = x_dist), 
              method = "glm", 
              method.args = list(family = binomial), 
              se = FALSE, 
              size = 1) +

  # Theme
  theme_light() +
  scale_color_manual(values = rcartocolor::carto_pal(4, "Safe")[c(4,1,3,2)]) +
  scale_size_manual(values = c(1, 2, 3, 4)) +
  xlab(expression(log[e] (Effect_size))) +
  ylab("Power") +
  labs(col = "", linetype = "Test") +
  facet_grid(type ~ conv_or_not) +
  theme(legend.position = "bottom")
```

@@

Shape of the non-linearity pattern is another factor contributed to the simulation model. Figure \ref{fig:poly-power-j}

@@

heter_conv_sim %>%
  filter(x_dist == "uniform") %>%
  mutate(log_effect_size = log(effect_size)) %>%
  mutate(reject = as.numeric(BP_p_value < 0.05)) %>%
  mutate(a = c("Left-triangle", "Butterfly", "Right-triangle")[a + 2]) %>%
  mutate(conv_or_not = "Conventional") %>%
  bind_rows(vi_survey %>%
  filter(!attention_check) %>%
  filter(!null_lineup) %>%
  filter(x_dist == "uniform", type == "heteroskedasticity") %>%
  mutate(log_effect_size = log(effect_size)) %>%
  mutate(reject = as.numeric(p_value < 0.05)) %>%
  mutate(a = c("Left-triangle", "Butterfly", "Right-triangle")[a + 2]) %>%
  mutate(conv_or_not = "Visual")) %>%
  mutate(a = factor(a, levels = c("Left-triangle", "Butterfly", "Right-triangle"))) %>%
  select(a, log_effect_size, reject, conv_or_not) %>%
  ggplot() +
  geom_smooth(aes(log_effect_size, reject, col = a), 
              se = FALSE, 
              method = "glm",
              method.args = list(family = binomial())) +
  xlab(expression(log[e] (Effect_size))) +
  ylab("Power") +
  scale_color_manual(values = rcartocolor::carto_pal(3, "Vivid")) +
  theme_light() +
  theme(legend.position = "bottom") +
  facet_wrap(~conv_or_not) +
  labs(col = "Heteroskedasticity shape")
```
